Chromosomal duplication can be an evolutionary detour on the path to adaptation

### 2022-09-18 ms/ms.tex
MP small comments

The model includes the effects of natural selection, genetic drift, aneuploidy, and mutation, and follows a population of cells characterized by both their genotype: 
euploid wild-type, \euwt, is the ancestral diploid genotype; 
euploid mutant, \eumt, has a diploid karyotype and a single beneficial mutation; 
aneuploid wild-type, \anwt, has an extra chromosome due to a chromosome duplication event; and
aneuploid mutant, \anmt, has and extra chromosome and a beneficial mutation.
Fitness values of the different genotypes are denoted by $w_{\euwt}$, $w_{\eumt}$, $w_{\anwt}$, and $w_{\anmt}$, and the rate of mutation and aneuploidy are denoted by $\mu$ and $\delta$. See \Cref{fig:models} for an illustration of the model.

@@

% without back transitions:
The estimated mutation rate, $\mu=2.965\cdot10^{-6}\ [2.718\cdot10^{-7}-3.589\cdot10^{-6}]$, corresponds to a mutation target size of $\sim 10^{4}$, assuming the mutation rate per base pair is roughly $2\cdot10^{-10}$ (ref.~\citep{Zhu2014}) or $3.3\cdot10^{-10}$ (ref.~\citep{Lynch2008}).
The estimated aneuploidy rate, $\delta=1.72\cdot10^{-3}\ [1.47\cdot10^{-3}-2.786\cdot10^{-3}]$ is higher than in previous studies: for chromosome III in diploid \yeast, \citet{Zhu2014} estimated $6.7\cdot10^{-6}$ chromosome gain events per generation, and \citet{Kumaran2013} estimate $3.0-4.3\cdot10^{-5}$ chromosome loss events per generation (95\% confidence interval).
The estimated fitness values are $w_{\anwt}=1.022\ [1.021-1.023]$,
$w_{\anmt}=1.025\ [1.024-1.026]$,
$w_{\eumt}=1.028\ [1.026-1.029]$, all relative to the fitness of \euwt, which is set to $w_{\euwt}=1$. 
Thus, we can infer that the cost of trisomy is $c = w_{\eumt}-w_{\anmt} = 0.003$ (or 0.3\%) and the benefit of trisomy is $w_{\anwt}-1-c=0.019$ (1.9\%), whereas the benefit of the beneficial mutation is $w_{\eumt}-1=0.028$ (2.8\%).

@@

>>>>

The model includes the effects of natural selection, genetic drift, aneuploidy, and mutation, and follows a population of cells characterized by their genotype: 
euploid wild-type, \euwt, is the ancestral diploid genotype; 
euploid mutant, \eumt, has a diploid karyotype and a single beneficial mutation; 
aneuploid wild-type, \anwt, has an extra chromosome due to a chromosome duplication event; and
aneuploid mutant, \anmt, has and extra chromosome and a beneficial mutation.
Fitness values of the different genotypes are denoted by $w_{\euwt}$, $w_{\eumt}$, $w_{\anwt}$, and $w_{\anmt}$, and the rate of mutation and aneuploidy are denoted by $\mu$ and $\delta$. See \Cref{fig:models} for an illustration of the model.

@@

% without back transitions:
The estimated beneficial mutation rate, $\mu=2.965\cdot10^{-6}\ [2.718\cdot10^{-7}-3.589\cdot10^{-6}]$, corresponds to a mutation target size of $\sim 10^{4}$, assuming the mutation rate per base pair is roughly $2\cdot10^{-10}$ (ref.~\citep{Zhu2014}) or $3.3\cdot10^{-10}$ (ref.~\citep{Lynch2008}).
The estimated aneuploidy rate, $\delta=1.72\cdot10^{-3}\ [1.47\cdot10^{-3}-2.786\cdot10^{-3}]$ is higher than in previous studies: for chromosome III in diploid \yeast, \citet{Zhu2014} estimated $6.7\cdot10^{-6}$ chromosome gain events per generation, and \citet{Kumaran2013} estimate $3.0-4.3\cdot10^{-5}$ chromosome loss events per generation (95\% confidence interval).
The estimated fitness values are $w_{\anwt}=1.022\ [1.021-1.023]$,
$w_{\anmt}=1.025\ [1.024-1.026]$,
$w_{\eumt}=1.028\ [1.026-1.029]$, all relative to the fitness of \euwt, which is set to $w_{\euwt}=1$. 
Thus, we can infer that the cost of trisomy is $c = w_{\eumt}-w_{\anmt} = 0.003$ (or 0.3\%) and the benefit of trisomy is $w_{\anwt}-1-c=0.019$ (1.9\%), whereas the benefit of the beneficial mutation is $w_{\eumt}-1=0.028$ (2.8\%).

@@

% TODO caveats - MP
